---
title: |
  Public-sector employment in \
  Kern County, California
thanks: "Department of Economics, California State University, Bakersfield. Email: \\href{emailto:richryan@csub.edu}{richryan@csub.edu}."
date: October 30, 2025
date-format: long
author: 
  - name: Rich Ryan
    corresponding: true
    affiliations:
      - name: California State University, Bakersfield
format:
  pdf:
    documentclass: scrartcl
    toc: false
    number-sections: true
    colorlinks: true
    template-partials:
      - title.tex
# format:
#   docx:
#     toc: false
#     number-sections: true
#     highlight-style: github
#     colorlink: true
fig-cap-location: bottom
---

```{r}
#| echo: false
#| include: false
library(targets)
source("R/packages.R")

tar_load(plt_shares)

plt_shares_date_start <- min(plt_shares$dat$date)
plt_shares_date_end <- max(plt_shares$dat$date)
plt_shares_caption <- paste0("Shares of public-sector employment in Kern County, California, ", 
                           month(plt_shares_date_start, label = TRUE, abbr = FALSE), " ",
                           year(plt_shares_date_start), " to ", 
                           month(plt_shares_date_end, label = TRUE, abbr = FALSE), " ",
                           year(plt_shares_date_end))

plt_shares_end <- filter(plt_shares$dat, date == max(date))

tar_load(plt_wages)
dat_wages <- plt_wages$dat
plt_wages_date_start <- min(plt_wages$dat$date)
plt_wages_date_end <- max(plt_wages$dat$date)
dat_wages_end <- filter(dat_wages, date == max(date)) |> 
  ungroup() |> 
  mutate(federal_ratio = 100 * avg_wkly_wage / max(avg_wkly_wage * (own_title == "Federal Government")))

plt_wages_caption_1 <- paste0("Sectoral wages in Kern County, California, ", 
                           year(plt_wages_date_start), " quarter ",
                           quarter(plt_wages_date_start),
                           " to ", 
                           year(plt_wages_date_end), " quarter ",
                           quarter(plt_wages_date_end), "."
                           )
plt_wages_caption_2 <- "Deflated by personal-consumption expenditures price index."
plt_wages_caption <- paste(plt_wages_caption_1, plt_wages_caption_2)
```

# Introduction

On September 30, 2025, the
director of the Office of Management and Budget,
Russell Vought,
sent a [memorandum](https://www.whitehouse.gov/wp-content/uploads/2025/09/M-25-35-Status-of-Agency-Operations.pdf) to the heads of executive departments and agencies within the US government.
The memorandum informed the heads that
appropriations for the fiscal year ending at 11:59 p.m. that evening would expire.
<!-- which would trigger a government shutdown. -->
Affected agencies were instructed to "execute their plans for an orderly shutdown."

Appropriations are laws of Congress that grant agencies the power to govern money.
When those powers expired at midnight,
agencies were required to furlough many federal workers.
While these workers will be paid retroactively in accordance with a 2019 law,
the absence of paychecks and continued interest in reducing the size of the federal workforce
will affect the economy in Kern County, California.
<!-- the shutdown will affect the economy in Kern County, California. -->

To partially understand the effect the shutdown will have on Kern's economy,
I investigate how employment is allocated across the public and private sectors using
<!-- composition of employment using  -->
data from the Quarterly Census of Employment and Wages.
<!-- In brief, -->
A main message is that, in Kern County,

* out of every 100 jobs, 
`r round(filter(plt_shares_end, own_title == "Federal Government") |> pull(share), 0)` are part of the federal government, whereas
`r round(100 - sum(plt_shares_end$share), 0)` are in the private sector; and
* federal employees earn, on average, each week
$`r prettyNum(round(filter(dat_wages_end, own_title == "Federal Government") |> pull(avg_wkly_wage), 0), big.mark = ",")`.


The remainder of this article describes how these statistics were computed.
Replication materials, including data on private-sector employment and public-sector employment at the federal, state, and local levels,
<!-- the data used to construct figures [-@fig-empl] and [-@fig-shares], -->
are available at 
[https://github.com/richryan/kern-public-sector-employment](https://github.com/richryan/kern-public-sector-employment).

<!-- What effect will the shutdown have on Kern's economy? -->
<!-- The answer to that question is impossible to know. -->
<!-- The answer to a related question, however, may give a hint at the effect. -->
<!-- Using data from the -->


<!-- These workers were sent home without pay, -->
<!-- prevented from even volunteering their services to -->
<!-- ensure Congress retains the power of the purse. -->

<!-- What does this mean for the economy in Kern County, California? -->
<!-- While the answer to this question is unknowable, -->
<!-- the answer to a related question  -->

# Public-sector employment in Kern County, California

The Quarterly Census of Employment and Wages program of the US Bureau of Labor Statistics 
covers "[more than 95 percent of U.S. jobs](https://www.bls.gov/opub/hom/cew/),"
which makes the data suitable to investigate employment at the county level.
The employment data are published at a monthly frequency and are available by ownership type.
Employment is the count of filled jobs,
regardless of 
whether the position is full or part time or
whether the position is temporary or permanent.[^seasadj]

[^seasadj]: All the time series are seasonally adjusted using the US Census Bureau’s X-13ARIMA-SETS procedure.

```{r}
#| include: false
tar_load(plt_empl)
plt_empl_date_start <- min(plt_empl$dat$date)
plt_empl_date_end <- max(plt_empl$dat$date)
plt_empl_caption <- paste0("Public-sector employment in Kern County, California, ", 
                           month(plt_empl_date_start, label = TRUE, abbr = FALSE), " ",
                           year(plt_empl_date_start), " to ", 
                           month(plt_empl_date_end, label = TRUE, abbr = FALSE), " ",
                           year(plt_empl_date_end))

# Levels
plt_empl_private_end <- filter(plt_empl$dat, date == max(date), own_title == "Private") |> pull(empl)
plt_empl_public_end <- filter(plt_empl$dat, date == max(date), own_title != "Private") |> 
  ungroup() |>
  summarise(empl = sum(empl))
plt_empl_jobs <- filter(plt_empl$dat, date == max(date))

print_jobs <- function(x) {
  prettyNum(round(x, digits = 0), big.mark = ",")
}
```


Employment across the private and public sectors is shown in figure [-@fig-empl].
Public-sector employment includes government workers at the
federal, state, and local levels.
Figure [-@fig-empl] indicates that most jobs are in the private sector.
In `r month(plt_empl_date_end, label = TRUE, abbr = FALSE)` `r year(plt_empl_date_end)`, there were
`r prettyNum(round(plt_empl_private_end, 0), big.mark = ",")` jobs in the private sector and 
`r prettyNum(round(plt_empl_public_end, 0), big.mark = ",")` jobs in the public sector.

The majority of public-sector jobs are in local government.
These include
[jobs](https://www.bls.gov/spotlight/2021/occupational-employment-and-wages-in-state-and-local-government/home.htm) 
held by teachers and workers who provide administrative support.
In `r month(plt_empl_date_end, label = TRUE, abbr = FALSE)` `r year(plt_empl_date_end)`, 
public-sector employment comprised
`r print_jobs(filter(plt_empl_jobs, own_title == "Local Government") |> pull(empl))` jobs in local government,
`r print_jobs(filter(plt_empl_jobs, own_title == "Federal Government") |> pull(empl))` jobs in federal government, and
`r print_jobs(filter(plt_empl_jobs, own_title == "State Government") |> pull(empl))` jobs in state government.

```{r}
#| echo: false
#| label: fig-empl
#| fig-cap: !expr plt_empl_caption
plt_empl$plt
```

Just `r round(filter(plt_shares_end, own_title == "Federal Government") |> pull(share), 1)` percent of jobs
are held by federal workers.
This statistic can be seen in figure [-@fig-shares],
which plots public-sector employment as a percentage of total employment.
In contrast to federal jobs,
`r round(filter(plt_shares_end, own_title == "Local Government") |> pull(share), 1)` percent of jobs are at the local level and
`r round(filter(plt_shares_end, own_title == "State Government") |> pull(share), 1)` percent of jobs are at the state level.

```{r}
#| echo: false
#| label: fig-shares
#| fig-cap: !expr plt_shares_caption
plt_shares$plt
```

# Wages across the public and private sectors in Kern County, California

In addition to data on employment,
the QCEW program also publishes data on wages.
These data are available at a quarterly frequency.
Wages refer to [total compensation](https://www.bls.gov/opub/hom/cew/),
"including bonuses, 
stock options, severance pay, profit distributions, 
the cash value of meals and lodging, tips and other gratuities, and, in some states, employer contributions to certain deferred compensation plans (such as 401(k) plans)."
The average weekly wage reports the ratio of total compensation paid during the quarter to average employment over the quarter 
divided by 13, for the 13 weeks in the quarter.

```{r}
#| echo: false
#| label: fig-wages
#| fig-cap: !expr plt_wages_caption
plt_wages$plt
```

Average weekly wages by sector are reported in figure [-@fig-wages].


# Conclusion

Across Kern each month,
workers earn paychecks by producing goods and structures or providing services.
The goods, services, and structures are purchased 
using funds that come out of workers' paychecks.
This circular flow is threatened by the government shutdown.
Even though just 
`r round(filter(plt_shares_end, own_title == "Federal Government") |> pull(share), 1)` 
percent of workers  are employed by the federal government in Kern,
the complex linkages  that connect various forms of economic activity may magnify the 
postponement of income allocated to federal workers.

