---
title: "PROJECT: kern-public-sector-employment"
author: "Rich Ryan"
email: "richryan@csub.edu"
format: gfm
---

## Public-sector employment in Kern County, California

How many public-sector jobs are there in Kern County, California?
This question is answered using data from the Quarterly Census of Employment and Wages.

## Posit Cloud and GitHub

The project was completed using Posit Cloud.
If you are interested in interacting with GitHub using Posit Cloud,
then these notes may be useful to you.

Here is the procedure I used to interact with GitHub through Posit Cloud:

* Create a [fine-grained personal access token](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/managing-your-personal-access-tokens#fine-grained-personal-access-tokens).
  * In the upper-right corner of any page on GitHub, click your profile photo, then click **Settings**
  * In the left sidebar, click **Developer** settings (at the bottom).
  * In the left sidebar, under **Personal access tokens**, click **Fine-grained tokens**.
  * Click **Generate new token**.
  * Under **Token name**, enter a name for the token.
  * Under **Expiration**, select an expiration for the token.
  * Optionally, under **Description**, add a note to describe the purpose of the token.
* Saved the personal access token using `gitcreds::gitcreds_set()`.
* Check the configuration of credentials: `git config -- list`.
Confirm that credentials are stored.
(And perhaps run `git config --global credential.helper store`.)

These steps should let Posit Cloud push and pull from GitHub.
  
## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
